Modelling last glacial cycle ice dynamics in the Alps

### 2017-01-25 articles/cycle/alps-cycle.tex
Add notes and ideas for co-autors.
Also fix wrong EPICA scaling factore in records table.

@@

    \begin{itemize}
    \item We conducted a 5/,km-resolution sensitivity study.
    \item We used three palaeo-temperature records (Table~\ref{tab:records}).
    \item All records scaled to fit the LGM extent (Table~\ref{tab:records}).
    \item All yield glaciation during MIS~4 and 2 (Fig.~\ref{fig:timeseries}).
    \item But otherwise very different results (Fig.~\ref{fig:timeseries}).
    \item Two yield two large MIS~4 ice extent (Fig.~\ref{fig:footprints}).
    \item Only EPICA yield realistic MIS~4 (Fig.~\ref{fig:footprints}).
    \end{itemize}

@@

    \begin{figure}
      \centerline{\includegraphics{alpcyc_hr_timing}}
      \caption{%
        \textbf{(a)} Timing of the Last Glacial Maximum (LGM) given by the
        modelled age (colour mapping) and value (200 m contours) of maximum
        surface elevation throughout the entire simulation.
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model (black curve),
        and modelled glaciated area around the LGM (coloured curve). The LGM
        is here modelled as a time-transgressive event.}
      \label{fig:timing}
    \end{figure}

    \begin{figure}
      \centerline{\includegraphics{alpcyc_hr_profiles}}
      \caption{%
        Modelled extent of glaciation along selected profiles for the Lyon,
        Solothurn, Rhine and Ivrea glaciers.
        \todo{Implement extraction of ice thickness along a more detailed
              profile, and add the other glaciers.}}
      \label{fig:profiles}
    \end{figure}

@@

    \begin{table*}
      \caption{%
        Palaeo-temperature proxy records and scaling factors yielding
        temperature offset time-series used to force the ice sheet model
        through the last glacial cycle (Fig.~\ref{fig:timeseries}). $f$
        corresponds to the scaling factor adopted to yield Last Glacial Maximum
        ice limits in the vicinity of mapped end moraines
        (Fig.~\ref{fig:footprints}a), and $[{\Delta}T_{\textrm{TS}}]_{32}^{22}$
        refers to the resulting mean temperature anomaly during the period 32
        to~22\,\unit{ka} after scaling.}
      \label{tab:records}
      \noindent\small\makebox[\textwidth]
      {\begin{tabular}{lccccccl}
        \toprule

@@

        EPICA     &  75{\degree}06$^{\prime}$\,S   % 75.1
                  & 123{\degree}21$^{\prime}$\,E   % 123.35
                  & 3233
                  & \chem{\delta^{18}O}
                  & 0.46 & $-$9.5  % -9.2055
                  & \citet{Jouzel.etal.2007} \\

@@

    \begin{figure}
      \centerline{\includegraphics{alpcyc_hr_inputs}}
      \caption{%
        \textbf{(a)} July mean near-surface air temperature and
        \textbf{(b)} January precipitation from WorldClim
        \citep[1960--1990]{Hijmans.etal.2005}, and
        \textbf{(c)} Mordern July standard deviation of daily mean temperature
        from the ERA-Interim \citep[1979--2012]{Dee.etal.2011} from the
        reference monthly climatology used to force the surface mass balance
        (PDD) component of the ice sheet model.
        \textbf{(d)} Geothermal heat flow from applying the similarity method
        to multiple geophysical proxies \citep{Goutorbe.etal.2011} used as a
        boundary condition to the bedrock thermal model 3\,km below the
        ice-bedrock interface.}
      \label{fig:inputs}
    \end{figure}

>>>>

    \note{For the Nature Geoscience format this abstract will need to be
          converted to an ``introductory paragraph''.}

@@

    \begin{itemize}
    \item We conducted a 5\,km-resolution sensitivity study.
    \item We used three palaeo-temperature records (Table~\ref{tab:records}).
    \item All records scaled to fit the LGM extent (Table~\ref{tab:records}).
    \item All yield glaciation during MIS~4 and 2 (Fig.~\ref{fig:timeseries}).
    \item But otherwise very different results (Fig.~\ref{fig:timeseries}).
    \item Two yield two large MIS~4 ice extent (Fig.~\ref{fig:footprints}).
    \item Only EPICA yield realistic MIS~4 (Fig.~\ref{fig:footprints}).
    \end{itemize}

    \note{An important assumption in the method is that the MIS~4 glaciation
          was less (or equally) extensive than the MIS~2, i.e. that the LGM
          outline from \citet{Ehlers.etal.2011} dates indeed from the LGM
          (MIS~2). A different assumption could be that this outline represents
          the most extensive glaciation in the last 120\,ka. I we allow the
          MIS~2 glaciation to be a bit smaller than MIS~4, simulations driven
          by GRIP and MD01-2444 also give reasonable results.}

@@

    \begin{figure}
      \centerline{\includegraphics{alpcyc_hr_timing}}
      \caption{%
        \textbf{(a)} Timing of the Last Glacial Maximum (LGM) given by the
        modelled age (colour mapping) and value (200 m contours) of maximum
        surface elevation throughout the entire simulation.
        \textbf{(b)} Temperature offset time-series from the EPICA ice core
        used as palaeo-climate forcing for the ice flow model (black curve),
        and modelled glaciated area around the LGM (coloured curve). The LGM
        is here modelled as a time-transgressive event.
        \note{The Nature Geoscience format is limited to four figures. Now we
              indeed have four figures, albeit large ones.
              Figure~\ref{fig:timing} will interest the dating people but is
              not crucial for the paper. My plan is to include in the main text
              for the first submission and be prepared that the editor may want
              to move it into supplementary material.}}
      \label{fig:timing}
    \end{figure}

    \begin{figure}
      \centerline{\includegraphics{alpcyc_hr_profiles}}
      \caption{%
        Modelled extent of glaciation along selected profiles for the Lyon,
        Solothurn, Rhine and Ivrea glaciers.
        \todo{Implement extraction of ice thickness along a more detailed
              profile, and add the other glaciers.}
        \idea{We can also move this as panel b on Fig.~\ref{fig:timing}, with
              profiles drawn on panel a. But then we should probably limit it
              to two glaciers instead of four.}}
      \label{fig:profiles}
    \end{figure}

@@

    \begin{table*}
      \caption{%
        Palaeo-temperature proxy records and scaling factors yielding
        temperature offset time-series used to force the ice sheet model
        through the last glacial cycle (Fig.~\ref{fig:timeseries}). $f$
        corresponds to the scaling factor adopted to yield Last Glacial Maximum
        ice limits in the vicinity of mapped end moraines
        (Fig.~\ref{fig:footprints}a), and $[{\Delta}T_{\textrm{TS}}]_{32}^{22}$
        refers to the resulting mean temperature anomaly during the period 32
        to~22\,\unit{ka} after scaling.
        \note{The EPICA scaling factor is close to one, and that ended up
              yielding too much ice in the high-resolution simulation
              (Fig.~\ref{fig:lgmvel}), though not so at low resotion
              (Fig.~\ref{fig:footprints}b). If I re-run the whole thing, I will
              probably fix it to one (i.e. use EPICA temperature anomalies
              without applying any scaling).}}
      \label{tab:records}
      \noindent\small\makebox[\textwidth]
      {\begin{tabular}{lccccccl}
        \toprule

@@

        EPICA     &  75{\degree}06$^{\prime}$\,S   % 75.1
                  & 123{\degree}21$^{\prime}$\,E   % 123.35
                  & 3233
                  & \chem{\delta^{18}O}
                  & 1.03 & $-$9.5  % -9.2055
                  & \citet{Jouzel.etal.2007} \\

@@

    \begin{figure}
      \centerline{\includegraphics{alpcyc_hr_inputs}}
      \caption{%
        \textbf{(a)} July mean near-surface air temperature and
        \textbf{(b)} January precipitation from WorldClim
        \citep[1960--1990]{Hijmans.etal.2005}, and
        \textbf{(c)} Mordern July standard deviation of daily mean temperature
        from the ERA-Interim \citep[1979--2012]{Dee.etal.2011} from the
        reference monthly climatology used to force the surface mass balance
        (PDD) component of the ice sheet model.
        \textbf{(d)} Geothermal heat flow from applying the similarity method
        to multiple geophysical proxies \citep{Goutorbe.etal.2011} used as a
        boundary condition to the bedrock thermal model 3\,km below the
        ice-bedrock interface.
        \idea{Since this will be in the supplement, I can also add additional
              information such as winter temperature, summer precipitation,
              winter PDD SD, and/or basal topography with modern glaciers
              removed, e.g. in the Bernese Alps.}}
      \label{fig:inputs}
    \end{figure}
